---
title: "Rtweet Hwk"
output: html_notebook
---

#Q1 load libraries and load the data
```{r}
library(tidyverse)
library(janitor)

tweets <- read_csv("data/code_clan_tweets.csv") %>% clean_names()

names(tweets)
dim(tweets)
```
#Q2 get total number of likes for "original" (unquoted) tweets

```{r}
tweets %>% 
  filter(is_quote == FALSE) %>% 
  summarise(total_favorites = sum(favorite_count))

#425 total favourites for original tweets
```

#Q3 get mean retweets, excl quoted tweets, grouped by source

```{r}
tweets %>% 
  filter(is_quote == FALSE) %>% 
  group_by(source) %>% 
  summarise(mean_retweets = mean(retweet_count)) %>% 
  ungroup()
```

#Q4 summarise total favourites, grouped by media type, ordered descending
#assume missing media type = text

```{r}
tweets %>% 
  mutate(media_type = if_else(
    is.na(media_type),
    "text",
    media_type
  )) %>% 
  group_by(media_type) %>% 
  summarise(total_favorites = sum(favorite_count)) %>% 
  arrange(desc(total_favorites)) %>% 
  ungroup()
```
#Q5 find mean length of CC tweet both using the existing column and by checking
#the string length
#str_length doesn't work because it gets stuck on unrecognised characters!
#this returns NA values in char_count
#try str_count instead - this works
#we now have two different mean values, seemingly because the display_text_width
#discounts any @twitter_handles

```{r}
tweets %>% 
  filter(is_quote == FALSE) %>% 
  mutate(char_count = str_count(text), .after = display_text_width) %>% 
  summarise(across(display_text_width:char_count, ~ mean(.)))
```

#Q6 join code_clan_info to add a url column to tweets

```{r}
code_clan_info <- read_csv("data/code_clan_info.csv") %>% clean_names()

#tweet_id looks like a good key to use - check it is unique in each dataset

tweets %>%
  distinct(tweet_id) %>% 
  count() == count(tweets)

code_clan_info %>% 
  distinct(tweet_id) %>% 
  count() == count(code_clan_info)

#tweet_id seems to be unique, so let's proceed with the task
#first get the columns we want from code_clan_info (tweet_id and status_url)


code_clan_info %>% 
  select(1:2) %>% 
  right_join(tweets, by = c("tweet_id" = "tweet_id"))

#edit - try this with a nested select.  Avoids a right join!
#put this in a new tibble for Q7

codeclan_hashtags <- tweets %>% 
  left_join(
    select(code_clan_info, 1:2),
    by = c("tweet_id" = "tweet_id")
  )

```

#Q7
#get just the tweet_id and the hashtags columns
#convert hashtags to lower case
#keep only records with hashtags

```{r}
codeclan_hashtags <-codeclan_hashtags %>% 
  select(tweet_id, hashtags) %>% 
  filter(!is.na(hashtags)) %>% 
  mutate(hashtags = str_to_lower(hashtags))

#this has worked startingly well
```

#Q8 find tweets where hashtags starts with c(
#not quite sure why the question mentions str_sub just yet...
#had to use \\( to escape for the regex

```{r}
codeclan_hashtags %>% 
  filter(str_detect(hashtags, "c\\("))
```

#Q9 find tweets mentioning "Edinburgh" and count them
#sticking to the question rigidly here - exact matches for "Edinburgh" only
#33 tweets matching

```{r}
tweets %>% 
  filter(str_detect(text, "Edinburgh")) %>% 
  count()
```

#read the hint... trying again with case insensitive match
#same result

```{r}
tweets %>% 
  filter(str_detect(text, "(?i)Edinburgh")) %>% 
  count()
```

#Q10 get the twitter handles
#going to assume "@" only appears at the start of handles, in tweets...
#this works, but the resulting data is somewhat messy

```{r}
str_split(tweets$text, " ") %>%
  flatten_chr() %>% 
  tibble() %>% 
  filter(str_detect(., "@"))
```

#OK... have googled and found that twitter restricts handles to 15 characters,
#alphanumeric or underscores only
#so, let's try a regex approach

```{r}
str_extract_all(tweets$text, "@[a-zA-Z_]{1,15}") %>% 
  flatten_chr()
```





