---
title: "Simple Linear Regression"
output: html_notebook
---

```{r}
library(tidyverse)
library(modelr)
library(ggfortify)
```

```{r}
pm <- read_csv("data/project_management.csv")
```

```{r}
pm %>% 
  ggplot() +
  aes(x = estimated_length, y = actual_length) +
  geom_point() +
  geom_text(aes(label = X1))
```

```{r}
cor(pm$estimated_length, pm$actual_length)
```
#Very strong correlation
#Therefore a successful regression analysis is likely to be possible

```{r}
my_model <- lm(actual_length ~ estimated_length, data = pm)

autoplot(my_model)
```
```{r}
my_model

glance_output <- glance(my_model)

glance_output$r.squared
```
#1.2 week increase in actual length per 1 week increase in estimated length
#estimated_length explains 65% of the variation in actual_length

```{r}
autoplot(my_model)

summary(my_model)
```
#residuals/fitted and scale/location plots indicate that model might not be reliable
#however p-value is very close to zero
#inconclusive

#res-lev
#5 and 18 are outliers; 5 influential and 18 non-influential

```{r}
autoplot(my_model, which = 4)
```

```{r}
pm_omit <- pm %>% filter(X1 != 5)
my3_model <- lm(actual_length ~ estimated_length, data = pm_omit)
summary(my3_model)
```

```{r}
pm %>% 
  ggplot() +
  aes(x = estimated_length, y = actual_length) +
  geom_point() +
  geom_abline(intercept = my_model[["coefficients"]][["(Intercept)"]],
              slope = my_model[["coefficients"]][["estimated_length"]],
              col = "red"
  ) +
  geom_abline(intercept = my2_model[["coefficients"]][["(Intercept)"]],
              slope = my2_model[["coefficients"]][["estimated_length"]],
              col = "green"
  ) +
  geom_abline(intercept = my3_model[["coefficients"]][["(Intercept)"]],
              slope = my3_model[["coefficients"]][["estimated_length"]],
              col = "blue"
  )
  
```

