---
title: "ggplot Homework 2"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(CodeClanData)
```

```{r}
theme <- 
  theme(
    text = element_text(size = 12),
    title = element_text(size = 14),
    axis.text = element_text(size = 10),
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "grey90", linetype = "dashed")
  )

colour_scheme <- c('#7fc97f','#beaed4','#fdc086','#ffff99','#386cb0','#f0027f','#bf5b17','#666666')
```

#MVP1
#Data is all numeric/factor so should be read to analyse
```{r}
backpack <- backpack
```

#MVP2
#Mostly around 10kg with an outlier at 25kg, main distribution is from 2-20
```{r}
backpack %>% 
  ggplot() +
  aes(x = backpack_weight) +
  geom_histogram(binwidth = 2, col = "white") +
  theme +
  labs(
    x = "Backpack Weight",
    y = "Frequency",
    title = "Frequency Distribution of Backpack Weights\n"
  )
```

#MVP3
#There are no obvious differences between weights for each sex
```{r}
backpack %>% 
  ggplot() +
  aes(x = backpack_weight, fill = sex) +
  geom_histogram(binwidth = 2, col = "white", position = "dodge") +
  theme +
  labs(
    x = "Backpack Weight",
    y = "Frequency",
    title = "Frequency Distribution of Backpack Weights\n",
    subtitle = "Split by sex"
  )

backpack %>% 
  ggplot() +
  aes(y = backpack_weight, x = sex) +
  geom_point() +
  theme +
  labs(
    x = "Backpack Weight",
    y = "Sex",
    title = "Backpack Weights by Sex\n"
  )
```

#MVP4
#There is a slight positive correlation between backpack weight and body weight
#But not an obvious one
```{r}
backpack %>% 
  ggplot() +
  aes(x = body_weight, y = backpack_weight) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme +
  labs(
    x = "Body Weight",
    y = "Backpack Weight",
    title = "Backpack Weight vs Body Weight\n"
  )
```

#MVP5
#There are no major differences between sex
#Females show a less clear correlation
```{r}
backpack %>% 
  ggplot() +
  aes(x = body_weight, y = backpack_weight) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ sex, scales = "free") +
  theme +
  labs(
    x = "Body Weight",
    y = "Backpack Weight",
    title = "Backpack Weight vs Body Weight",
    subtitle = "NB scales differ between plots\n"
  )
```

#MVP6
Roughly one third of students report back pain
```{r}
backpack %>% 
  mutate(back_problems = ifelse(back_problems == 1, "Yes", "No")) %>% 
  ggplot() +
  aes(x = back_problems) +
  geom_bar() +
  geom_text(
    stat = "count",
    aes(label = stat(count), vjust = 2, col = "white"),
    show.legend = FALSE
    ) +
  scale_color_identity() +
  theme +
  labs(
    x = "Back Problems",
    y = "Count",
    title = "Back Problems Amongst Students"
  )
```

#MVP7
#Yes, nearly half of females report back pain vs about one sixth of males
```{r}
backpack %>% 
  mutate(back_problems = ifelse(back_problems == 1, "Yes", "No")) %>% 
  ggplot() +
  aes(x = back_problems) +
  geom_bar() +
  geom_text(
    stat = "count",
    aes(label = stat(count), vjust = 2, col = "white"),
    show.legend = FALSE
    ) +
  facet_wrap(~sex) +
  scale_color_identity() +
  theme +
  labs(
    x = "Back Problems",
    y = "Count",
    title = "Back Problems Amongst Students"
  )
```

#MVP8

```{r}
backpack %>%
  mutate(back_problems = ifelse(back_problems == 1, "Yes", "No")) %>%
  ggplot() +
  aes(x = body_weight, y = backpack_weight) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(back_problems ~ sex, scales = "fr") +
  theme +
  labs(
    x = "Body Weight",
    y = "Backpack Weight",
    title = "Backpack Weight vs Body Weight",
    subtitle = "Split by Sex and Back Problems\n"
  )

backpack %>% 
  mutate(back_problems = ifelse(back_problems == 1, "Yes", "No")) %>%
  mutate(bw_quartile = as_factor(cut_number(
      x = body_weight,
      n = 4, 
      labels = c("Quart1", "Quart2", "Quart3", "Quart4")
      ))) %>% 
  ggplot() +
  aes(x = back_problems) +
  geom_bar() +
  facet_grid(bw_quartile ~ sex) +
  scale_color_identity() +
  theme +
  labs(
    x = "Back Problems",
    y = "Count",
    title = "Back Problems Amongst Students",
    subtitle = "Split by Sex and Body Weight Quartile"
  )
```

